---
title: "the interactive spatial plots of Spatial-Gaussian model"
author: "Xieqing Yu"
date: "2024-12-23"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This R Markdown document is made interactive using Shiny. Unlike the more traditional workflow of creating static reports, you can now create documents that allow your readers to change the assumptions underlying your analysis and see the results immediately. 

To learn more, see [Interactive Documents](http://rmarkdown.rstudio.com/authoring_shiny.html).

## Inputs and Outputs

You can embed Shiny inputs and outputs in your document. Outputs are automatically updated whenever inputs change.  This demonstrates how a standard R plot can be made interactive by wrapping it in the Shiny `renderPlot` function. The `selectInput` and `sliderInput` functions create the input widgets used to drive the plot.

```{r}
library(shiny)
library(leaflet)
library(sf)
library(dplyr)
library(lubridate)

# Load shapefile
california_counties <- st_read("/Users/xieqingyu/Desktop/ca_counties/rsconnect/shinyapps.io/tracyyu2024/CA_Counties.shp") %>%
  rename(county = NAME) %>%  # Rename NAME to county
  st_transform(crs = 4326)   # Ensurede CRS is WGS 84 for leaflet

# Load dataset
data <- read.csv("/Users/xieqingyu/Desktop/ca_counties/rsconnect/shinyapps.io/tracyyu2024/updated_dataset_with_predictions.csv")

# Ensure date columns are in Date format
data$date <- as.Date(data$date, format = "%Y-%m-%d")
data$prediction_date <- as.Date(data$prediction_date, format = "%Y-%m-%d")

# Define UI
ui <- fluidPage(
  titlePanel("Spatial-temporal Gaussian Regression Model"),
  sidebarLayout(
    sidebarPanel(
     sliderInput(
  "selected_date",
  "Select Date:",
  min = min(c(data$date, data$prediction_date), na.rm = TRUE),
  max = max(c(data$date, data$prediction_date), na.rm = TRUE),
  value = min(data$date, na.rm = TRUE),
  timeFormat = "%Y-%m-%d",
  step = 1
    )
    ),
    mainPanel(
      leafletOutput("true_cases_map"),
      leafletOutput("predicted_cases_map")
    )
  )
)

# Define Server
server <- function(input, output) {
  # Ensure column names are lowercase
  california_counties <- california_counties %>% rename_all(tolower)
  data <- data %>% rename_all(tolower)

  # Ensure county columns are character
  california_counties$county <- as.character(california_counties$county)
  data$county <- as.character(data$county)

  # Remove rows with missing county or numeric values
  california_counties <- california_counties %>% filter(!is.na(county))
  data <- data %>%
    filter(!is.na(county) & !is.na(daily_change_per_100k) & !is.na(predicted_daily_change_per_100k))

  # Filter data based on the selected date
  filtered_data <- reactive({
    data %>%
      filter(date == input$selected_date | prediction_date == input$selected_date)
  })

  # Merge filtered data with county shapefile for true cases
  merged_true <- reactive({
    california_counties %>%
      left_join(
        filtered_data() %>%
          filter(date == input$selected_date),
        by = "county"
      )
  })

  # Merge filtered data with county shapefile for predicted cases
  merged_predicted <- reactive({
    california_counties %>%
      left_join(
        filtered_data() %>%
          filter(prediction_date == input$selected_date),
        by = "county"
      )
  })

  # Render true cases map
  output$true_cases_map <- renderLeaflet({
    true_data <- merged_true()
    if (nrow(true_data) == 0 || all(is.na(true_data$daily_change_per_100k))) {
      return(NULL)  # Do not render if no valid data
    }
    leaflet(true_data) %>%
      addTiles() %>%
      addPolygons(
        fillColor = ~colorNumeric("YlOrRd", daily_change_per_100k, na.color = "transparent")(daily_change_per_100k),
        color = "black",
        weight = 1,
        fillOpacity = 0.7,
        label = ~paste0(county, ": ", round(daily_change_per_100k, 2), " per 100k")
      ) %>%
      addLegend(
        pal = colorNumeric("YlOrRd", range(true_data$daily_change_per_100k, na.rm = TRUE), na.color = "transparent"),
        values = true_data$daily_change_per_100k,
        title = "True Daily Cases per 100k",
        position = "bottomright"
      )
  })

  # Render predicted cases map
  output$predicted_cases_map <- renderLeaflet({
    predicted_data <- merged_predicted()
    if (nrow(predicted_data) == 0 || all(is.na(predicted_data$predicted_daily_change_per_100k))) {
      return(NULL)  # Do not render if no valid data
    }
    leaflet(predicted_data) %>%
      addTiles() %>%
      addPolygons(
        fillColor = ~colorNumeric("YlOrRd", predicted_daily_change_per_100k, na.color = "transparent")(predicted_daily_change_per_100k),
        color = "black",
        weight = 1,
        fillOpacity = 0.7,
        label = ~paste0(county, ": ", round(predicted_daily_change_per_100k, 2), " per 100k")
      ) %>%
      addLegend(
        pal = colorNumeric("YlOrRd", range(predicted_data$predicted_daily_change_per_100k, na.rm = TRUE), na.color = "transparent"),
        values = predicted_data$predicted_daily_change_per_100k,
        title = "Predicted Daily Cases per 100k",
        position = "bottomright"
      )
  })
}

shinyApp(ui, server)


```